---
title: "Reproducible Research: Peer Assessment 1"
author: "jtdata"
date: "01/17/2015"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}

rawActivity <- read.csv("activity.csv", header=TRUE)
activity <- rawActivity[which(rawActivity$steps != "NA"), ]
```

## What is mean total number of steps taken per day?
```{r}
library(plyr) 
dailyTotal <- ddply(activity, .(date), summarise, steps=sum(steps))
hist(dailyTotal$steps, main="Number of Steps per Day", 
     xlab="Total Steps per Day", col="blue")
mean(dailyTotal$steps)
median(dailyTotal$steps)
```

## What is the average daily activity pattern?
```{r}
intervalAverage <- ddply(activity, .(interval), summarise, steps=mean(steps))
plot(intervalAverage$interval, intervalAverage$steps, type="l", col="red",
     xlab="5 Minute Interval", ylab="Number of Steps (Average)")
# Maximum average steps in a 5 miute interval:
intervalAverage[intervalAverage$steps==max(intervalAverage$steps), ]
colnames(intervalAverage)[2] <- "intervalAvg"
```

## Imputing missing values
```{r}
# Total missing values:
sum(is.na(activity$steps))
# Fill NA's with surrounding 5 minute average:
fullActivity <- arrange(join(activity, intervalAverage), interval)
fullActivity$steps[is.na(fullActivity$steps)] <- fullActivity$intervalAvg[is.na(fullActivity$steps)]
# Plot results:
newDailyTotal <- ddply(fullActivity, .(date), summarise, steps=sum(steps))
hist(newDailyTotal$steps, main="Number of Steps",
     xlab="Total Daily Steps", col="blue", )
# New Mean and Median Values:
mean(newDailyTotal$steps)
median(newDailyTotal$steps)
totalSteps1 <- sum(activity$steps)
totalSteps2 <- sum(fullActivity$steps)
diff <- totalSteps1 - totalSteps2
```
Mean value did not change because the 5 minute average was used to fill in missing values. The histogram illustrates the difference in total steps, and median changed since adding values will shift the middle value.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lattice)
weekdays <- weekdays(as.Date(fullActivity$date))
fullActivityWD <- transform(fullActivity, day=weekdays)
fullActivityWD$wk <- ifelse(fullActivityWD$day %in% c("Saturday", "Sunday"), "weekend", "weekday")
intervalAverageWD <- ddply(fullActivityWD, .(interval, wk), summarise, steps=mean(steps))
xyplot(steps ~ interval | wk, data=intervalAverageWD, layout=c(1,2), type="l")
```


